Developing a preprint culture in biology

### 2012-09-28 ms.tex
Proofreading; demonic physics-enrichment programme
Gave this a thorough proof-reading for grammar (subject-verb and plural
agreement).  I also clarified and rearranged some sentences.  I
reinserted my idea about reaching physicists through submitting to ArXiv,
this time as a two-sentence comment.  This is a big problem in
evolutionary biology and it would be good to have more about it in
print.  I hope we can find some way to keep it.

\documentclass[letterpaper,twocolumn,superscriptaddress,showkeys]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\hypersetup
{
    colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

@@

Public preprints servers allow authors to make their manuscripts publicly
available before, or in parallel to, submitting them to journals for traditional
peer-review. The idea behind preprint servers is fundamentally simple: to make
the results available to the scientific community as soon as possible rather
than wait until the peer-review process is fully completed. The point of arXiv
and open preprints servers is not to avoid the peer-review process. Almost all
manuscripts on arXiv are submitted to peer-review.  The point is that sharing
ideas as quickly as possible using preprint servers has numerous advantages.
These include rapid dissemination of work in progress to a wider audience, not
only scientists and the general public but also readers in developing countries
where access to subscription journals are often a limiting factor. Furthermore,
preprints provide the opportunity to solicit feedback from a larger pool of
reviewers and can be seen as an integral aspect of a vigorous peer-review
process \cite{hoc12}. Thus, preprint servers increase the number of
opportunities for review and revision prior to publication, resulting in higher
quality submissions which could also alleviate reviewer burden.

The idea gained popularity 20 years ago with the advent of arXiv, an open
preprint server widely used in physics and mathematics \cite{gin11}.  It is also
part of other fields' culture. Paul Krugman noted that, in economics, the
\emph{traditional model of submit, get refereed, publish, and then people will
read your work broke down a long time ago. In fact, it had more or less fallen
apart by the early 80s} \cite{kru12}. In addition to a section in arXiv,
economists have also the RePEc (Research Papers in Economics) initiative, which
aims to create an archive of working papers, manuscripts, and book chapters.
Yet, despite the success of the approach in other fields of inquiry, most
manuscripts in biology are never submitted to any public archive.  In this
article, we will first highlight the advantages of open preprints servers for
both scientists and publishers. We will also debunk a few misconceptions,
discuss the policies of major publishers in biology, and briefly review the most
popular open preprint servers.

@@

The first and most often discussed advantage of arXiv and open preprints is
speed \ref{fig:map}. The time between submission and the official publication of
a manuscript can be measured in months, sometime in years. For all this time,
the research is only known to a select few: colleagues, editors, reviewers.
Thus, the science cannot be used, discussed, or reviewed by the wider scientific
community. It is a problem for both scientists and publishers. Manuscripts that
are unknown cannot be integrated and thus take more time to be cited. It has
been shown that high-energy physics, with its high arXiv submission rate, had
the highest immediacy among physics and mathematics \cite{pra05}.

Furthermore, the review process as a whole is critically over-loaded, because
the number of active scientists increases, because the pressure to publish
increases, and because of an effect dubbed ``the tragedy of the reviewers
commons'' REF.  At the same times, rejection rates are high in most journals
(REF), and when not invited to submit a revision, authors must start the whole
process all over again. It's thus no surprise that different initiatives emerged
over the last few years to decrease the time spent in review. XXX et coll.
(REF) called for the recycling and reuse of peer-reviews: by attaching previous
reviews and detailed replies to a new submission, both the editor and the
referees can gauge the work done on the manuscript, and perhaps evaluate it with
less prejudice. In a similar way, the \emph{Peerage of Science} initiative
allows authors to seek anonymous pre-review by their peers. Some journals now
accept to publish papers which received good evaluations, with \emph{Animal
Biology} having recently accepted first a paper reviewed entirely with the
\emph{Peerage of Science} \cite{abb12}, effectively outsourcing the review
process. A widespread use of preprint servers can achieve the same goal of
reducing the time spent in review. By putting a manuscript for open comments and
criticisms, the authors will receive valuable feedback and can improve the
version which will be submitted. With a rich enough community of scientists
depositing preprints, and commenting on them, the process of an open pre-review
can become widespread and will overall increase the quality of first
submissions.

@@

Public preprint servers offer a much fairer way to establish intellectual
priority by making the work available when done. Some manuscripts will spend
much more time in the review process than others. Surprisingly, there is a
perception in biology that public preprints make it easier to steal ideas, as if
a scientific idea was only complete once published in a peer-reviewed journal.
Mathematicians and physicists have embraced arXiv in part to establish priority
in a fair way and biologists should do so too, or at least not be afraid to
submit to make their research public for fear of been snooped \cite{cal12}.

Prepublication reviews by a small network of colleagues is an important part of
the scientific process, which is attested by the fact that nearly all published
papers acknowledge comments by people not listed as co-authors.  Preprint
servers simply offer a way to extend this network of colleagues to the entire
scientific community. It ensures that science is not constrained by small
networks of scientists exchanging ideas.  Paul Ginsparg created arXiv.org in
part for democratic reasons: he wanted everyone from students in small
universities to Ivy-League professors to have access to the most recent
scientific \emph{ideas}.  Ginsparg's revolutionary idea was simply to use the
power of the internet for preprints, not just for the end product, so the
scientific process can be open as soon as possible.

@@

Submitting to a preprint server is becoming more common in biology, even though
it still involves a minority of papers. The quantitative biology section in
arXiv is experiencing faster growth in submissions than any other field
\cite{cal12}. Examples include a recent series of papers on the theory of
natural selection that was posted to arXiv simultaneously with its publication
in the \emph{Journal of Evolutionary Biology}
\cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}.

Most scientific journals are preprint-friendly: Nature, PLOS, BMC, PNAS, Science
(mostly) \ref{table:policies}, and all the journals from Elsevier and Springer.
Very recently, the Ecological Society of America and the Genetics Society of
America changed their policies to allow public preprints. Few scientific
publications will not consider a manuscript submitted to a public preprint
server.  Still, a few journals adopt a ``by default'' hostile attitude towards
preprints, mostly due to the lack of clear policy of the publishers, or perhaps
because a preprint culture has not developed in biology and the practice is
still considered unusual. As an example, Wiley-Blackwell, which publishes some
leading journals, has no official policy on the subject \ref{table:policies}.

@@

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.  It
provides a reliable citation system for all eprints and is especially popular in
high-energy physics. Physicist Paul Ginsparg created arXiv in 1991 for
theoretical high-energy physicists to communicate preprints via email and ftp,
and soon thereafter adopted the newly created world-wide
web\cite{jackson2002preprints}.  arXiv now receives over 7,000 submissions per
month (\url{http://arxiv.org/show_monthly_submissions}) and divides its
submissions into subcategories of physics, mathematics, computer science,
quantitative biology, finance and statistics.  The quantitative biology category
includes subcategories for Populations and Evolution, Quantitative Methods and
other categories that may be of interest to biologists.

@@

% To review: 
To submit to arXiv, the authors must either have their work available under an
open license or grant arXiv a non-exclusive and irrevocable license to
distribute the article. In either case, arXiv does not require copyright
transfer and only requires the rights to distribute submitted articles in
perpetuity. Thus, submitting to arXiv does not in itself prevent the authors
from transferring their rights to a publisher, which is why most publishers
tolerate arXiv, even though they ask the authors to transfer their right to them
upon acceptance of the article.

Most papers posted to arXiv are eventually printed in journals but
there are notable exceptions, such as Perelman's landmark paper
leading to the proof of the Poincar\'{e} conjecture
\cite{2002math.....11159P}. Yet, arXiv has never sought to replace
scientific journals and explicitly states that it serves a different
function as ``an openly accessible, moderated repository for scholarly
articles in specific scientific disciplines.'' arXiv is now
administered by the Cornell University Libraries, with funding coming
from voluntary pledges by academic institutions along with matching
funds from the Simons Foundation \cite{arxiv_future}.  One-hundred
twenty six of the top two-hundred institutions in terms of downloads
have provided the operating budget for arXiv over the next five years.
This plan reduces the financial burden on Cornell University and
transfers governance to a collaborative community in accordance with
arXiv's key principles.  arXiv takes numerous measures to ensure that
the repository will remain permanently available and submissions will
be readable.

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. All figshare content has a unique digital object identifier (DOI) like
any journal article, thus offering a permanent and stable link to the content.
A flexible tag system is used to classify each item. All content can be
commented and is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0. The CC-BY license grant the rights to
\emph{copy, distribute, display and perform the work and make derivative works
based on it only if they give the author or licensor the credits in the manner
specified by these.} The CC0 is the most permissive license and make the work
public domain (no rights reserved) or, if it is not possible in the given
jurisdiction, provides a simple permissive license.

One of the biggest advantage of figshare over arXiv is that is it not limited to
quantitative sciences. arXiv.org has sections on quantitative biology but might
not be appropriate for non-quantitative work. With its flexible approach to
preprints, figshare offer an important alternative to arXiv for empirical
biologists. Furthermore, by allowing all types of contents, figshare arguably
provides an archive for early results (e.g.: figures, lab presentations).

@@

Open preprints servers offer a great opportunity for open science, especially if
the community embrace the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work. These initiatives are
important to fully exploit the potential of open preprints servers. Posting
preprints online increases the community of available informal peer reviewers,
and uses the internet for its original community-building purposes.  Preprint
servers also facilitate communication between disciplines, bridging cultural as
well as geographic divides. The advantages are clear and the costs are low.

>>>>

\documentclass[letterpaper,twocolumn,superscriptaddress,showkeys]{revtex4}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\hypersetup
{
    colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

@@

Public preprint servers allow authors to make manuscripts publicly
available before, or in parallel to, submitting them to journals for
traditional peer-review. The rationale for preprint servers is
fundamentally simple: to make the results available to the scientific
community as soon as possible rather than to wait until the
peer-review process is fully completed. The goal of arXiv and open
preprint servers is not to avoid the peer-review process; almost all
manuscripts on arXiv are submitted to peer-review.  Sharing ideas as
quickly as possible using preprint servers has numerous advantages.
These include rapid dissemination of work in progress to a wider
audience, including readers in developing countries where access to
subscription journals is often a limiting factor in scientific
research.  Furthermore, preprints provide the opportunity to solicit
feedback from a larger pool of reviewers and can be seen as an
integral aspect of a vigorous peer-review process \cite{hoc12}.
Preprint servers therefore increase the number of opportunities for
review and revision prior to publication, resulting in higher quality
submissions and also could alleviate reviewer burden.

Preprints gained popularity 20 years ago with the advent of arXiv, an
open preprint server widely used in physics and mathematics
\cite{gin11}. Preprints are also integral to the culture of other
scientific fields.  Paul Krugman noted that, in economics, the
\emph{traditional model of submit, get refereed, publish, and then
  people will read your work broke down a long time ago. In fact, it
  had more or less fallen apart by the early 80s} \cite{kru12}. In
addition to a section in arXiv, economists have also the RePEc
(Research Papers in Economics) initiative, which aims to create an
archive of working papers, manuscripts, and book chapters.  Despite
the success of this approach in other fields, most manuscripts in
biology are never submitted to any public archive.
% JJA: This is a tenuous statement: MANY if not most biomedical labs
% in the US are funded by NIH, which requires submission to Pubmed
% Central, which many would consider a preprint archive.  In other
% words, a biomedical editor might take issue with this statement,
% especially if he considers "biology" to be just his part of biology
% --- I know plenty of people who only think of "biology" as
% biomedical science
In this article, we will first highlight the advantages of open
preprint servers for both scientists and publishers. We will also
debunk a few misconceptions, discuss the policies of major publishers
in biology, and briefly review the most popular open preprint servers.

@@

The first and most often discussed advantage of arXiv and open
preprints is speed (Figure~\ref{fig:map}). The time between submission
and the official publication of a manuscript can be measured in
months, sometime in years. For all this time, the research is known
only to a select few: colleagues, editors, reviewers.  Thus, the
science cannot be used, discussed, or reviewed by the wider scientific
community.  This stage of manuscript preparation is a problem for both
scientists and publishers.  Manuscripts that are unknown cannot be
used and thus take more time to be cited. It has been shown that
high-energy physics, with its high ArXiv submission rate, had the
highest immediacy among physics and mathematics \cite{pra05}.
% JJA: Can you say what "immediacy" means?

Furthermore, the review process as a whole is critically over-loaded,
because as the number of active scientists increases, because the
pressure to publish increases, and because of an effect dubbed ``the
tragedy of the reviewers commons'' REF.  At the same time, rejection
rates are high in most journals (REF), and when not invited to submit
a revision, authors must start the whole process all over again.
Initiatives to reduce time from submission to publication have emerged
across the scientific community XXX et coll.  (REF) called for the
recycling and reuse of peer-reviews: by attaching previous reviews and
detailed replies to a new submission, both the editor and the referees
can gauge the work done on the manuscript, and perhaps evaluate it
with less prejudice. In a similar way, the \emph{Peerage of Science}
initiative allows authors to seek anonymous pre-review by their
peers. Some journals now accept to publish papers which received good
evaluations, with \emph{Animal Biology} having recently accepted first
a paper reviewed entirely with the \emph{Peerage of Science}
\cite{abb12}, effectively outsourcing the review process. A widespread
use of preprint servers can achieve the same goal of reducing the time
spent in review. By putting a manuscript for open comments and
criticisms, the authors will receive valuable feedback and can improve
the version which will be submitted. With a rich enough community of
scientists depositing preprints, and commenting on them, the process
of an open pre-review can become widespread and will overall increase
the quality of first submissions.

@@

Public preprint servers offer a much fairer way to establish
intellectual priority by making the work available when done. Some
manuscripts will spend much more time in the review process than
others. Surprisingly, there is a perception in biology that public
preprints make it easier to steal ideas, as if a scientific idea was
only complete once published in a peer-reviewed journal.
Mathematicians and physicists have embraced arXiv in part to establish
priority in a fair way \cite{cal12}.
% JJA: I snipped the part about getting "snooped" (do you mean
% "scooped"?) as it seems normative and might annoy some readers.  The
% "should" should be implicit.

Prepublication reviews by a small network of colleagues is an
important part of the scientific process, attested by the fact that
nearly all published papers acknowledge comments by people not listed
as co-authors.  Preprint servers simply offer a way to extend this
network of colleagues to the entire scientific community. It ensures
that science is not constrained by small networks of scientists
exchanging ideas.  Paul Ginsparg created arXiv.org in part for
democratic reasons: he wanted everyone from students in small
universities to Ivy-League professors to have access to the most
recent scientific \emph{ideas}.  Ginsparg's revolutionary idea was
simply to use the power of the internet for preprints, not just for
the end product, so the scientific process can be open as soon as
possible.

@@

Submitting to a preprint server is becoming more common in biology,
even though it still involves a minority of papers. The quantitative
biology section in arXiv is experiencing faster growth in submissions
than any other field \cite{cal12}.  Examples include a recent series
of papers on the theory of natural selection that was posted to ArXiv
simultaneously with its publication in the \emph{Journal of
  Evolutionary Biology}
\cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}.  Other
submissions in this category include evolutionary and ecological
theory by authors trained in physics and computer science.  Since
authors in these fields regularly check ArXiv, submitting preprints
may be the most effective way biologists can help others avoid
``repeated work'' \cite{de2011contribution} and form a synthetic
community of evolutionary theorists from disparate backgrounds.

Most scientific journals are preprint-friendly: Nature, PLOS, BMC,
PNAS, Science (mostly) \ref{table:policies}, and all the journals from
Elsevier and Springer.  The Ecological Society of America and the
Genetics Society of America recently changed their policies to allow
public preprints.  Few scientific publications will not consider a
manuscript submitted to a public preprint server.  Still, a few
journals adopt a ``by default'' hostile attitude towards preprints,
mostly due to the lack of clear policy of the publishers, or perhaps
because a preprint culture has not developed in biology and the
practice is still considered unusual. As an example, Wiley-Blackwell,
which publishes some leading journals, has no official policy on the
subject \ref{table:policies}.

@@

% JJA: It's not crucial.  My main point was that there are certain
% non-obvious cultural reasons that arXiv submissions are common.
% It's not just because arXiv started in theoretical physics.  It's
% also that people who use TeX tend to have a certain attitude, I
% would argue inherited from Don Knuth, and people who use TeX and
% LaTeX _talk to each other_.  So perhaps this was a polite way of
% saying that people in math and physics are more community-oriented,
% and people in biology tend to walk around paranoid about getting
% scooped.  Some people also might think that they can't submit to
% arXiv because they use Word.

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.
% Ni!!!!  I really don't like starting sentences with "it"
arXiv provides a reliable citation system for all eprints and is
especially popular in high-energy physics. Physicist Paul Ginsparg
created arXiv in 1991 for theoretical high-energy physicists to
communicate preprints via email and ftp, and soon thereafter adopted
the newly created world-wide web\cite{jackson2002preprints}.  ArXiv
now receives over 7,000 submissions per month
(\url{http://arxiv.org/show_monthly_submissions}) and divides its
submissions into subcategories of physics, mathematics, computer
science, quantitative biology, finance and statistics.  The
quantitative biology category includes subcategories for Populations
and Evolution, Quantitative Methods and other categories that may be
of interest to biologists.

@@

% To review: 
Authors must either have their arXiv submission available under an
open license or grant arXiv a non-exclusive and irrevocable license to
distribute the work. In either case, arXiv does not require copyright
transfer and only requires the rights to distribute submitted articles
in perpetuity. Thus, submitting to arXiv does not in itself prevent
the authors from transferring their rights to a publisher, which is
why most publishers tolerate arXiv, even though they ask the authors
to transfer their right to them upon acceptance of the article.

Most papers posted to arXiv are eventually printed in journals but
there are notable exceptions, such as Perelman's landmark paper
leading to the proof of the Poincar\'{e} conjecture
\cite{2002math.....11159P}.  However, arXiv has never sought to
replace scientific journals and explicitly states that it serves a
different function as ``an openly accessible, moderated repository for
scholarly articles in specific scientific disciplines.'' arXiv is now
administered by the Cornell University Libraries, with funding coming
from voluntary pledges by academic institutions along with matching
funds from the Simons Foundation \cite{arxiv_future}.  One-hundred
twenty six of the top two-hundred institutions in terms of downloads
have provided the operating budget for arXiv over the next five years.
This plan reduces the financial burden on Cornell University and
transfers governance to a collaborative community in accordance with
arXiv's key principles.  arXiv takes numerous measures to ensure that
the repository will remain permanently available and submissions will
be readable.

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an
open server that allows scientists to submit any research output:
manuscript, figures, datasets, videos, theses, presentations, and so
on. There are no rules to limit what constitutes a research output
and, unlike arXiv, there is no endorser system. All figshare content
has a unique digital object identifier (DOI) like any journal article,
thus offering a permanent and stable link to the content.  A flexible
tag system is used to classify each item. All content can be commented
and is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0. The CC-BY license grants
rights to \emph{copy, distribute, display and perform the work and
  make derivative works based on it only if they give the author or
  licensor the credits in the manner specified by these.}  CC0 is the
most permissive license and effectively puts the work in public domain
(no rights reserved) or, if it is not possible in the given
jurisdiction, provides a simple permissive license.

One of the biggest advantage of figshare over arXiv is that is it not
limited to quantitative sciences. arXiv.org has sections on
quantitative biology but might not be appropriate for non-quantitative
work. With its flexible approach to preprints, figshare offers an
important alternative to arXiv for empirical biologists. Furthermore,
by allowing all types of content, figshare arguably provides an
archive for early results (e.g.: figures, lab presentations).

@@

Open preprint servers offer a great opportunity for open science,
especially if the community embraces the idea of discussing
preprints. Initiatives like Haldane's Sieve
(\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a new
blog discussing arXiv papers in population genetics, will help make
arXiv attractive for scientists looking to promote their work. These
initiatives are important to fully exploit the potential of open
preprints servers. Posting preprints online increases the community of
available informal peer reviewers, and uses the internet for its
original community-building purposes.  Preprint servers also
facilitate communication between disciplines, bridging cultural as
well as geographic divides. The advantages are clear and the costs are
low.
